PLAE enables quick and easy searching across one million ocular transcriptomes

### 2022-07-14 manuscript_00.Rmd
test tweaks

```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}
#knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
knitr::opts_chunk$set(out.width = '100%', dpi=300)
library(tidyverse)
library(citr)
library(cowplot)
library(ggrepel)
library(colorspace)
library(flextable)
library(captioner)
library(glue)
library(tictoc)
#tic()()
data_dir_vGiga <- '/Volumes/McGaughey_S/data/scEiaD/'
data_dir_vPLAE <- '/Volumes/McGaughey_S/data/scEiaD_2022_02/'
#data_dir_vGiga <- '~/data/scEiaD_gigascience/'
#data_dir_vPLAE <- '~/data/scEiaD_2022_02/'
# setup caption-ing
fig_cap <- captioner("Figure")
supFig_cap <- captioner("Supplemental Figure")
tab_cap <- captioner("Table")
supTab_cap <- captioner("Supplemental Table")

@@

Across the cell types of the eye, we find `r diff_tab %>% nrow()` gene - celltype combinations that are significantly differentially expressed. We define significant as a log2 fold change greater than two as well as a p-value (corrected) less than $1 x 10^-5. We find a mean of 
`r diff_tab_sup$`Homo sapiens` %>% mean() %>% round()`, 
`r diff_tab_sup$`Mus musculus` %>% mean() %>% round()`, and 
`r diff_tab_sup$`Macaca fascicularis` %>% mean() %>% round()` differentially expressed genes in human, macaque, and mouse respectively (`r supTab_cap(name = 'diff_tab_sup_tab', display = 'cite')`). These variable results directly reflect the differing number of independent studies - mouse and human have substantially more than macaque (`r supTab_cap(name = 'well_supported_celltypes_tab', display = 'cite')`). The full differential table is made available as supplementary file "differential_expression_table.csv.gz." While this differential expression table is large, it can be mined further to give more precise results.

@@

Cell type marker genes can be proposed on the basis of existing knowledge about their function and/or based on high expression differences when comparing to other cell types. We can further narrow down the list of markers by using the  diversity of studies and organisms in scEiaD to propose a set of community supported cell type markers. We first identify a set of well supported cell types which we define as cell types which have two or more independent studies across both human and mouse (`r supTab_cap(name = 'well_supported_celltypes_tab', display = 'cite')`). These cell types were then assessed to identify differentially expressed genes which met the following criteria: 1. padj less than $1 x 10^-4
in both human and mouse, 2. mean log2 fold change greater than two, and 3. mean padj less than $1 x 10^-5

@@

```{r}
# tic()
# ws_oct <- ws_table %>% filter(`Well Supported` == 'Yes') %>% pull(`CellType (Predict)`)
# 
# consist_diff <- scEiaD_2020_v01 %>% 
#   tbl("diff_testing") %>% 
#   filter(Base %in% ws_oct) %>%  
#   filter(Group == 'CellType (Predict)', 
#          Against == 'All') %>% 
#   group_by(Base, Gene) %>% 
#   summarise(`mean log2FC` = mean(log2FoldChange), `sig count` = sum(padj < 1e-4), `mean padj` = mean(padj)) %>% 
#   filter(`mean log2FC` > 2, `sig count` > 1, `mean padj` < 1e-5) %>%  
#   collect() %>% 
#   mutate(GENEID = str_extract(Gene, 'ENS\\w+')) %>% 
#   left_join(gene_tab, by = 'GENEID') %>% 
#   relocate(SYMBOL, GENEID) %>% 
#   dplyr::rename(Symbol = SYMBOL, ID = GENEID) %>% 
#   arrange(-`mean log2FC`)
# 
# # id diff genes only seen once across the cell types
# uniq_g <- consist_diff %>% group_by(Symbol) %>% count() %>% filter(n == 1)
consist_diff %>% group_by(Base) %>% summarise(Count = n())
consist_diff %>% filter(Symbol %in% uniq_g$Symbol) %>% group_by(Base) %>% summarise(Count = n())

consist_diff %>% filter(Symbol %in% uniq_g$Symbol) %>% group_by(Base) %>% summarise(Count = n()) %>% ggplot(aes(x=`Base`, y = `Count` )) + geom_bar(stat='identity', position = 'dodge') + cowplot::theme_cowplot() + coord_flip() + xlab("CellTye (Predict)")
#toc()
```

@@

diff_tab_sup %>% 
  flextable()

ws_table %>% 
  flextable()

@@

diff_tab_sup_cap %>% flextable()
```

>>>>

```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}
#knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
knitr::opts_chunk$set(out.width = '100%', dpi=300)
library(tidyverse)
library(citr)
library(cowplot)
library(ggrepel)
library(colorspace)
library(flextable)
library(captioner)
library(glue)
library(tictoc)
#tic()()
#data_dir_vGiga <- '/Volumes/McGaughey_S/data/scEiaD/'
#data_dir_vPLAE <- '/Volumes/McGaughey_S/data/scEiaD_2022_02/'
data_dir_vGiga <- '~/data/scEiaD_gigascience/'
data_dir_vPLAE <- '~/data/scEiaD_2022_02/'
# setup caption-ing
fig_cap <- captioner("Figure")
supFig_cap <- captioner("Supplemental Figure")
tab_cap <- captioner("Table")
supTab_cap <- captioner("Supplemental Table")

@@

Across the cell types of the eye, we find `r diff_tab %>% nrow()` gene - celltype combinations that are significantly differentially expressed. We define significant as a log2 fold change greater than two as well as a p-value (corrected) less than $1 x 10^-5#. We find a mean of 
`r diff_tab_sup$"Homo sapiens" %>% mean() %>% round()`, 
`r diff_tab_sup$"Mus musculus" %>% mean() %>% round()`, and 
`r diff_tab_sup$"Macaca fascicularis" %>% mean() %>% round()` differentially expressed genes in human, macaque, and mouse respectively (`r supTab_cap(name = 'diff_tab_sup_tab', display = 'cite')`). These variable results directly reflect the differing number of independent studies - mouse and human have substantially more than macaque (`r supTab_cap(name = 'well_supported_celltypes_tab', display = 'cite')`). The full differential table is made available as supplementary file "differential_expression_table.csv.gz." While this differential expression table is large, it can be mined further to give more precise results.

@@

Cell type marker genes can be proposed on the basis of existing knowledge about their function and/or based on high expression differences when comparing to other cell types. We can further narrow down the list of markers by using the  diversity of studies and organisms in scEiaD to propose a set of community supported cell type markers. We first identify a set of well supported cell types which we define as cell types which have two or more independent studies across both human mouse (`r supTab_cap(name = 'well_supported_celltypes_tab', display = 'cite')`). These cell types were then assessed to identify differentially expressed genes which met the following criteria: 1. padj less than $1 x 10^-4$
in both human and mouse, 2. mean log2 fold change greater than two, and 3. mean padj less than $1 x 10^-5$

@@

```{r}
# tic()
# ws_oct <- ws_table %>% filter(`Well Supported` == 'Yes') %>% pull(`CellType (Predict)`)
# 
consist_diff2 <- scEiaD_2020_v01 %>%
  tbl("diff_testing") %>%
  filter(Base %in% ws_oct) %>%
  filter(Group == 'CellType (Predict)',
         Against == 'All',
         Organism %in% c('Homo sapiens','Mus musculus')) %>%
  group_by(Base, Gene) %>%
  summarise(`mean log2FC` = mean(log2FoldChange), `sig count` = sum(padj < 1e-5), `mean padj` = mean(padj)) %>%
  filter(`mean log2FC` > 2, `sig count` > 1, `mean padj` < 1e-5) %>%
  collect() %>%
  mutate(GENEID = str_extract(Gene, 'ENS\\w+')) %>%
  left_join(gene_tab, by = 'GENEID') %>%
  relocate(SYMBOL, GENEID) %>%
  dplyr::rename(Symbol = SYMBOL, ID = GENEID) %>%
  arrange(-`mean log2FC`)

# # id diff genes only seen once across the cell types
# uniq_g <- consist_diff %>% group_by(Symbol) %>% count() %>% filter(n == 1)
consist_diff %>% group_by(Base) %>% summarise(Count = n())
consist_diff %>% filter(Symbol %in% uniq_g$Symbol) %>% group_by(Base) %>% summarise(Count = n())

consist_diff %>% filter(Symbol %in% uniq_g$Symbol) %>% group_by(Base) %>% summarise(Count = n()) %>% ggplot(aes(x=`Base`, y = `Count` )) + geom_bar(stat='identity', position = 'dodge') + cowplot::theme_cowplot() + coord_flip() + xlab("CellType (Predict)")
#toc()
```

@@

# diff_tab_sup %>% 
#   flextable()
# 
# ws_table %>% 
#   flextable()

@@

diff_tab_sup %>% flextable()
```
